---
title: "Results"

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(fig.pos = 'H')
knitr::opts_chunk$set(results = 'asis')
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
library(readxl)
library(tidyverse)
library(gtsummary)
library(dplyr)
library(ggplot2)
library(minpack.lm)
library(nlme)
library(splines)
library(cowplot)
```



## Pateints' demographic and clinical characteristics


```{r, message = FALSE, cache = TRUE, warning=FALSE, echo = FALSE, results='asis'}

# https://livefreeordichotomize.com/2017/08/08/how-to-make-an-rmarkdown-website/

final_dataset <- read_csv("Sempa_final_pull_with_results.csv") %>%
  mutate(vl = ifelse(`Viral Load at Draw`== '<40', '40', ifelse(`Viral Load at Draw`=='NULL', "", `Viral Load at Draw`))) %>%
  mutate(logvl = log(as.numeric(vl),10),
         time_on_trt = -1 * `days since tx start`,
         hiv_antibody_value = `Sedia LAg Odn screen`,
         RaceEthnicity = `Race/Ethnicity`) %>%
  arrange(subject_label_blinded, time_on_trt)

dt01 <- final_dataset %>%
  group_by(subject_label_blinded) %>%
  mutate(`days since ART start` = max(time_on_trt, na.rm = T)) %>%
  ungroup() %>%
  distinct(subject_label_blinded, .keep_all = T) %>%
  dplyr::select(Group, Subtype, BiologicalSex, RaceEthnicity, `days since ART start`) %>%
  tbl_summary(missing = "no") %>% 
  modify_header(label = "Variable") %>%
  bold_labels()
dt01

```

Majority 18 out of 19 patients were Males, with HIV sub-type B. They had a 1213 days (IQR: 1087,1352) on since the start of treatment.

## Ploting a patients' HIV viral load and HIV antibody response before and after Antiretroviral therapy intiation among early suppressors


```{r, message = FALSE, cache = TRUE, warning=FALSE, echo = FALSE, results='asis'}
model1 <- nlme::lme(fixed = logvl ~ Group + bs(time_on_trt, 3),
                    random = ~ 1 | subject_label_blinded,
                    data = final_dataset %>%
                      filter(time_on_trt>=0),
                    na.action = na.exclude, control = lmeControl(opt = "optim")
)
# summary(model1)

pred_vl <- predict(model1, newdata = final_dataset %>%
                     filter(time_on_trt>=0), level = 0:1)
after_ART <- final_dataset %>%
  mutate(subject_label_blinded = as.character(subject_label_blinded)) %>%
  filter(time_on_trt>=0) %>%
  dplyr::select(subject_label_blinded, logvl, time_on_trt) %>%
  left_join(pred_vl %>% 
              mutate(subject_label_blinded = as.character(subject_label_blinded)) %>% 
              select(subject_label_blinded, predict.subject_label_blinded), by = 'subject_label_blinded')

plot_grid(
ggplot(data = final_dataset[final_dataset$Group == 'early suppressions',],
       aes(x = time_on_trt, y = logvl)) +
  geom_line(aes(color = as.factor(subject_label_blinded), linetype = Group), size = 1.5) +
  geom_smooth(data = after_ART %>% filter(subject_label_blinded!='74498805'), 
              aes(x = time_on_trt, y = predict.subject_label_blinded), size = 2,
              method = lm, formula = y ~ splines::bs(x, 3)) +
  theme(
    text = element_text(size = 20),
    plot.title = element_text(hjust = 0.5),
    axis.line = element_line(colour = "black"),
    axis.title = element_text(size = 18),
    panel.background = element_blank(),
    panel.border = element_blank(),
    plot.margin=unit(c(0,0,0,0), "null"),
    legend.position = "none"
  ) +
  ylab(expression(paste(Log[10], ' Viral load'))) + xlab(''), #+
  # ggtitle('HIV viral load and HIV antibody response'),

ggplot(data = final_dataset[final_dataset$Group == 'early suppressions',],
       aes(x = time_on_trt, y = `Sedia LAg Odn screen`)) + #as.factor(subject_label_blinded)
  geom_line(aes(color = as.factor(subject_label_blinded), linetype = Group), size = 1.5) +
  geom_smooth(data = final_dataset[final_dataset$Group == 'early suppressions',] %>% 
                filter(subject_label_blinded!=74498805) %>%
                filter(time_on_trt>=0), 
              aes(x = time_on_trt, y = `Sedia LAg Odn screen`), size = 2,
              method = lm, formula = y ~ splines::bs(x, 3)) +
  theme(
    text = element_text(size = 20),
    plot.title = element_text(hjust = 0.5),
    axis.line = element_line(colour = "black"),
    axis.text = element_text(size = 18),
    axis.title = element_text(size = 18),
    panel.background = element_blank(),
    panel.border = element_blank(),
    plot.margin=unit(c(0,0,0,0), "null"),
    legend.position = "none"
  ) +
  ylab('Sedia LAg ODn') + xlab('Time from ART start (Days)'),
ncol = 1#,
# labels = 'AUTO',
# label_colour = "red"
) 

```

The thick blue line indicates a smoothed line for either viral load or HIV antibody response drawn using B-splines with 3 knots on time since patients were initiated on ART

# Viral load Antibody model


```{r, message = FALSE, cache = TRUE, warning=FALSE, echo = FALSE, results='asis'}
model2 <- nlme::lme(fixed = logvl ~ Group + hiv_antibody_value + bs(time_on_trt, 3),
                    random = ~ 1 | subject_label_blinded,
                    data = final_dataset%>%
                      filter(time_on_trt>=0), 
                    na.action = na.exclude
)
sjPlot::tab_model(model2, show.icc = FALSE, show.re.var = FALSE)

```



```{r image_grobs, echo=FALSE, fig.show = "hold", out.width = "20%", fig.align = "default"}
knitr::include_graphics(
  c("03-Red_EDCTP.jpg",
    "EU-flag.jpg")
)
```